Simulating Dwarf Galaxy Evolution with Individual Star Feedback and Chemical Evolution

### 2018-02-13 main.tex
edits to introduction

\documentclass[twocolumn]{aastex61}
\pdfoutput=1 %for arXiv submission
\usepackage{amsmath,amstext}
\usepackage[T1]{fontenc}
\usepackage{apjfonts} 
\usepackage[figure,figure*]{hypcap}
\usepackage{multirow}
\usepackage{mathtools}
\renewcommand*{\sectionautorefname}{Section} %for \autoref
\renewcommand*{\subsectionautorefname}{Section} %for \autoref

@@

\begin{abstract}
We have developed a new model for stellar feedback in order to simulate the chemodynamical evolution of isolated dwarf galaxies at high resolution in the adaptive mesh refinement hydrodynamics code \textsc{Enzo}. For the first time in galaxy scale simulations, we are capable of capturing detailed stellar feedback from individual stars including stellar winds from massive stars and AGB stars, photoelectric heating, Lyman-Werner dissociation, ionizing radiation tracked through an adaptive ray-tracing radiative transfer method, and core collapse and Type Ia supernovae. 
%This star-by-star method allows us to directly account for variations in stellar properties that govern their feedback properties (mass, lifetime, surface temperature, and metallicity).
In addition, we follow the evolution of star-by-star stellar yields for 15 metal tracer species: C, N, O, Na, Mg, Si, S, Ca, Mn, Fe, Ni, As, Sr, Y, and Ba. We describe these methods in detail, including the additional physics used to model a realistic three-phase interstellar medium. We present the first application of these methods on a 500~Myr simulation of an isolated dwarf galaxy with properties similar to a Local Group ultrafaint dwarf galaxy. We demonstrate that our physics and feedback model is capable of producing a dwarf galaxy whose evolution is consistent with observations in both the Kennicutt-Schmidt relationship and extended Schmidt relationship. Effective feedback drives significant metal enriched outflows, leading to low metal retention fractions consistent with observations. Our low-dust content galaxy exhibits a significant H$_2$ mass fraction, formed primarily through gas-phase H$^{-}$ reactions. We end with a discussion on how this work will impact our understanding of galactic chemical evolution on a species-by-species basis.
\end{abstract}

@@

Due to its complexity, the detailed interstellar medium (ISM) and chemical abundance properties of galaxies
%beyond total metallicity
are sensitive tests of the underlying physical processes that govern galaxy evolution. Examining these in more detail in galaxy scale simulations is an important and exciting new discriminator between models. There is a considerable body of work studying the chemodynamical evolution of galaxies using cosmological hydrodynamics simulations \citep[e.g.][]{Lia2002,KawataGibson2003,Kobayashi2004,Tornatore2004,Romeo2005,OppenheimerDave2008,Wiersma2009,Shen2010,MUGS2010,ErisSimulation,Brook2014,Snaith2015,Oppenheimer2016,OWLS,EAGLE,FIRE}. These simulations, coupled with additional attention to feedback processes, have made remarkable progress in reproducing global galaxy trends such as evolution of the mass-metallicity relationship \citep[e.g.][]{Obreja2014, Ma2016, Dave2017, Torrey2017} and more detailed quantities such as metallicity distribution functions (MDFs) and abundance evolution of individual species \citep{Marcolini2008,Revaz2009,Sawala2010,RevazJablonka2012,Jeon2017,Hirai2017} .

@@

Fully Lagrangian schemes, like native smooth particle hydrodynamics (SPH), are incapable of capturing mixing between chemically inhomogeneous particles, as necessary for chemical evolution. This must be handled through sub-grid models of turbulent metal diffusion \citep[e.g.][]{Shen2010, Shen2013,Brook2014,Su2017a}. Even more, there are many possible sub-grid models that are not applicable in every regime \citep[see ][]{Revaz2016}, though recent studies have shown great improvements in characterizing various methods \citep{Escala2017}. While mixing occurs in Eulerian codes without sub-grid models, numerical diffusion tends to result in over-mixing, requiring very high resolution or additional viscosity to reduce. Molecular diffusion is certainly not resolved in any galaxy-scale simulation with either method, requiring additional sub-grid models; this can be particularly important for understanding the initial pollution of otherwise pristine gas \citep[see ][ and references therein]{PanScannapiecoScalo2013,Sarmento2017}. Moreover, metal mixing efficiencies may vary species-by-species, as suggested by recent work \citep[e.g.][]{Cohen2013, Roederer2014, FrebelNorris2015, Hirai2017, Cote2017}. Mixing behavior is tied critically to the feedback source (stellar winds, supernovae, and possibly more exotic sources) 
%mm in which they are injected to
    that inject metals into 
the ISM with different energies and on different timescales; the observational effect of this is poorly understood, however. The variations in how different methods handle sub-grid metal injection and metal mixing schemes can lead to uncertainties in connecting models to observations and the fundamental physics that drives galaxy evolution. \textbf{Still not sure about this sentence - but I do need a way to cite mentioned papers:} Extending these models to a grid-based code is a valuable and complementary approach towards investigating galactic chemical evolution that has only been done recently \citep{Few2012,Few2014,Vorobyov2015}.

Although increasing physical resolution reduces 
%mm one's 
reliance on sub-grid physics, at high particle mass resolution ($M \lesssim 10^3$ M$_{\odot}$) standard schemes for modeling stars as simple stellar populations (SSPs) are no longer valid \citep[as studied in detail in ][]{Revaz2016}. Below 10$^4$ M$_{\odot}$, SSPs do not fully sample the IMF, and cannot be considered average representations of stellar populations. This is acutely problematic at star particle masses comparable to or below the mass of the most massive individual star (or $\sim 100$ M$_{\odot}$). Various approaches exist to address this issue \citep[e.g.][]{Kobayashi2000,WeidnerKroupa2004,Pflamm-AltenburgKroupa2006,RevazJablonka2012,Kroupa2013,Rosdahl2015,Su2017b}, but none are without caveats \citep{Revaz2016}, save for schemes which begin to track the star-by-star information within a given particle by directly sampling the IMF at formation time \citep[e.g.][]{Hu2017}. 
%mm Perhaps the most naive, 
   The most straightforward
solution is to remove the SSP formalism entirely and simply track stars as individual particles. However, the number of particles required to do so make this computationally prohibitive in all but the smallest of galaxies ($M_{*} \lesssim 10^{7} {\rm M_{\odot}})$, let alone in a cosmological simulation.
%In this work, we 
%mm follow stars as individual star particles, allowing us to place unprecedented detail on our
%     do this, allowing unprecedented detail in our
% accounting of stellar feedback processes and chemical enrichment.

We introduce a new method for studying galactic chemical evolution that follows stars as individual star particles implemented in the adaptive mesh refinement code \texttt{Enzo}, designed for high resolution simulations of isolated galaxies. The relative simplicity of idealized, isolated galaxy evolution simulations allows for a focused, first-principles approach to studying multi-channel feedback mechanisms. We follow recent work using low mass dwarf galaxies as laboratories to study in detail how feedback governs galaxy evolution \citep{Forbes2016,Hu2016,Hu2017}.
%For example, \citet{Forbes2016} found that photoelectric heating can be a significant source of feedback in small galaxies. \citet{Hu2016} and \citet{Hu2017} examined the relative importance of various feedback physics in regulating ISM properties of dwarf galaxies. 
Our work builds upon our current understanding of feedback and galactic chemodynamics while making three notable advancements: 1) direct star-by-star modeling, 2) stellar wind feedback of both massive stars and AGB stars, and 3) using an adaptive ray tracing method to follow stellar ionizing radiation.

Using star-by-star modeling, we are able to capture in more detail the stellar yields from individual stars released over their lifetime. In addition to better capturing how individual metal species enrich the ISM, this allows us to chemically tag individual stars. This opens an exciting new channel for testing models of galaxy evolution by leveraging current and ongoing observations probing the detailed chemical abundances of stars in the Milky Way and Local Group, such as APOGEE and APOGEE2 \citep{APOGEE2010,APOGEE}, the Gaia-ESO survey \citep{Gaia}, and GALAH \citep{GALAH}.

This paper is the first in a series examining in detail the role that the individual components of multi-channel stellar feedback play in galaxy dynamical and chemical evolution. In this paper we introduce a new method of feedback physics and galactic chemical enrichment in the cosmological, hydrodynamics code \texttt{Enzo}. For the first time in galaxy-scale simulations, we model star formation down to particles representing individual stars, while simultaneously capturing stellar feedback through stellar winds, core collapse and Type Ia supernovae, photoelectric heating, and ionizing radiation through ray-tracing radiative transfer. In addition, we couple to a self-consistent set of stellar yields %mm tables from the
to track yields from massive star stellar winds, AGB star winds, and supernovae. We describe each mode of our multi-channel feedback 
%mm implementation 
in detail below, and briefly discuss their implementation in an isolated dwarf galaxy simulation.

@@

\section{Methods}
\label{sec:methods}
In 
%mm brief, 
     summary, 
we produce high-resolution galaxy-scale simulations tracking stars not as SSP's, but as individual stars sampled from an assumed IMF.
This allows us to follow star-by-star variations in feedback physics and stellar yields in detail. To properly model the ISM, we track non-equilibrium primordial chemistry (including molecular hydrogen) using \texttt{Grackle} \citep{GrackleMethod}, with heating and approximate self-shielding from a metagalactic ultraviolet (UV) background. We 
%mm 
    assume collisional ionization equilibrium for all other elements in order to
use updated \texttt{Cloudy} metal line cooling tables consistent with our self-shielding approximation (see Appendix ~\ref{appendix:cooling}), and an updated observationally motivated dust model for the low metallicity regimes studied here (Z $\lesssim 0.1$ Z$_{\odot}$). Each star is assigned basic properties (surface gravity, effective temperature, radius, lifetime) from tabulated stellar evolution models, which inform how the stars deposit their feedback. We directly track ionizing radiation from massive stars using an adaptive ray tracing radiative transfer method that includes the effects of radiation pressure. In addition, we follow the optically thin non-ionizing UV radiation from these stars leading to photoelectric heating and Lyman-Werner dissociation of molecular hydrogen. We track the stellar wind feedback and supernovae from these stars, depositing individual metal yields from both. We include AGB wind feedback and yields for lower mass stars, and track these directly as Type Ia supernovae progenitors. We follow yields for 15 individual metal species (C, N, O, Na, Mg, Si, S, Ca, Mn, Fe, Ni, As, Sr, Y, and Ba), chemically tagging each star
%mm
   as it forms
with the associated local gas abundances for each species. In addition, we track a total metal density field which is the sum of all metals, including those not directly tracked. This field is used to inform the heating/cooling physics and determines the metallicity of each star. These methods are discussed in full detail below.

>>>>

%mm2 substantive changes marked with this comment tag, minor changes and wordsmithing done silently.  12-15 Feb
\documentclass[twocolumn]{aastex61}
\pdfoutput=1 %for arXiv submission
\usepackage{amsmath,amstext}
\usepackage[T1]{fontenc}
\usepackage{apjfonts} 
\usepackage[figure,figure*]{hypcap}
\usepackage{multirow}
\usepackage{mathtools}
\renewcommand*{\sectionautorefname}{Section} %for \autoref
\renewcommand*{\subsectionautorefname}{Section} %for \autoref

@@

\begin{abstract}
We have developed a new model for stellar feedback in order to simulate the chemodynamical evolution of isolated dwarf galaxies at high resolution in the adaptive mesh refinement hydrodynamics code \textsc{Enzo}. For the first time in galaxy scale simulations, we simulate detailed stellar feedback from individual stars including stellar winds from massive stars and AGB stars, photoelectric heating, Lyman-Werner dissociation, ionizing radiation tracked through an adaptive ray-tracing radiative transfer method, and core collapse and Type Ia supernovae. 
%This star-by-star method allows us to directly account for variations in stellar properties that govern their feedback properties (mass, lifetime, surface temperature, and metallicity).
We furthermore follow the star-by-star stellar yields using tracer fields for 15 metal species: C, N, O, Na, Mg, Si, S, Ca, Mn, Fe, Ni, As, Sr, Y, and Ba. We describe these methods in detail, including the additional physics used to model a realistic three-phase interstellar medium. We present the first application of these methods in a 500~Myr simulation of an isolated dwarf galaxy with properties similar to a Local Group ultrafaint dwarf galaxy. We demonstrate that our physics and feedback model is capable of producing a dwarf galaxy whose evolution is consistent with observations in both the Kennicutt-Schmidt relationship and extended Schmidt relationship. Effective feedback drives significantly metal-enriched outflows, leading to low metal retention fractions consistent with observations. Our low-dust content galaxy exhibits a significant H$_2$ mass fraction, formed primarily through gas-phase H$^{-}$ reactions. We end with a discussion on how this work impacts our understanding of galactic chemical evolution on a species-by-species basis.
\end{abstract}

@@

%mm2 Due to its complexity, the 
Detailed interstellar medium (ISM) and chemical abundance properties of galaxies
%beyond total metallicity
are sensitive tests of the underlying physical processes that govern galaxy evolution. Examining these in more detail in galaxy scale simulations is an important and exciting new discriminator between models. There is a considerable body of work studying the chemodynamical evolution of galaxies using cosmological hydrodynamics simulations \citep[e.g.][]{Lia2002,KawataGibson2003,Kobayashi2004,Tornatore2004,Romeo2005,OppenheimerDave2008,Wiersma2009,Shen2010,MUGS2010,ErisSimulation,Brook2014,Snaith2015,Oppenheimer2016,OWLS,EAGLE,FIRE}. These simulations, coupled with additional attention to feedback processes, have made remarkable progress in reproducing global galaxy trends such as evolution of the mass-metallicity relationship \citep[e.g.][]{Obreja2014, Ma2016, Dave2017, Torrey2017} and more detailed quantities such as metallicity distribution functions (MDFs) and abundance evolution of individual species \citep{Marcolini2008,Revaz2009,Sawala2010,RevazJablonka2012,Jeon2017,Hirai2017} .

@@

%mm2 I think there does need to be some introduction to this discussion:
   However, the bulk of this work has been done with fully Lagrangian
%mm2 Fully Lagrangian schemes, like native 
smoothed particle hydrodynamics (SPH), 
%mm2 are incapable of capturing
    with a few recent exceptions  \citep{Few2012,Few2014,Vorobyov2015}. In its original form, 
    SPH cannot model
mixing between chemically inhomogeneous particles, as necessary for chemical evolution. This must be handled through sub-grid models of turbulent metal diffusion \citep[e.g.][]{Shen2010, Shen2013,Brook2014,Su2017a}. Even more, there are many possible sub-grid models that are not applicable in every regime \citep[see ][]{Revaz2016}, though recent studies have shown great improvements in characterizing various methods \citep{Escala2017}. While mixing occurs in Eulerian codes even without sub-grid models, numerical diffusion tends to result in over-mixing, requiring very high resolution or additional viscosity to reduce. Molecular diffusion %mm2 
    or even turbulent mixing
is certainly not resolved in any galaxy-scale simulation with either method, requiring additional sub-grid models; this can be particularly important for understanding the initial pollution of otherwise pristine gas \citep[see ][ and references therein]{PanScannapiecoScalo2013,Sarmento2017}. Moreover, metal mixing efficiencies may vary species-by-species, as suggested by recent work \citep[e.g.][]{Cohen2013, Roederer2014, FrebelNorris2015, Hirai2017, Cote2017}. Mixing behavior is tied critically to the feedback source (stellar winds, supernovae, and possibly more exotic sources) 
%mm in which they are injected to
    that inject metals into 
%mm2
    different phases of 
the ISM with different energies and on different timescales; the observational effect of this is poorly understood, however. The variations in how different methods handle sub-grid metal injection and metal mixing schemes can lead to uncertainties in connecting models to observations and the fundamental physics that drives galaxy evolution.
%mm2 [cited above] \textbf{Still not sure about this sentence - but I do need a way to cite mentioned papers:} Extending these models to a grid-based code is a valuable and complementary approach towards investigating galactic chemical evolution that has only been done recently \citep{Few2012,Few2014,Vorobyov2015}.

%mm2 [broke up sentence]
Increasing physical resolution reduces reliance on sub-grid physics for mixing.  However, at high particle mass resolution ($M \lesssim 10^3$ M$_{\odot}$) standard schemes for modeling stars as simple stellar populations (SSPs) then lose validity  \citep[as studied in detail by][]{Revaz2016}. Below 10$^4$ M$_{\odot}$, SSPs do not fully sample the initial mass function (IMF), and cannot be considered average representations of stellar populations. This is acutely problematic at star particle masses comparable to or below the mass of the most massive individual star ($\sim 100$ M$_{\odot}$). Various approaches exist to address this issue \citep[e.g.][]{Kobayashi2000,WeidnerKroupa2004,Pflamm-AltenburgKroupa2006,RevazJablonka2012,Kroupa2013,Rosdahl2015,Su2017b}, but none are without caveats \citep{Revaz2016}, save for schemes which begin to track the star-by-star information within a given particle by directly sampling the IMF at formation time \citep[e.g.][]{Hu2017}.  The most straightforward
solution is to remove the SSP formalism entirely and simply track stars as individual particles,
%mm2 [maybe it would be best to just remove this.] However, the number of particles required to do so make this computationally prohibitive in all but the smallest of galaxies ($M_{*} \lesssim 10^{7} {\rm M_{\odot}})$, let alone in a cosmological simulation.
    which we do here.

%mm2 [if we were to keep the previous sentence, we'd need to segue like this.]
%   Nevertheless, we 
We introduce a new method for studying galactic chemical evolution that follows stars as individual star particles implemented in the adaptive mesh refinement code \texttt{Enzo}, designed for high resolution simulations of isolated galaxies. The relative simplicity of idealized, isolated galaxy evolution simulations allows for a focused, first-principles approach to studying multi-channel feedback mechanisms. We follow recent work using low mass dwarf galaxies as laboratories to study in detail how feedback governs galaxy evolution \citep{Forbes2016,Hu2016,Hu2017}.
%For example, \citet{Forbes2016} found that photoelectric heating can be a significant source of feedback in small galaxies. \citet{Hu2016} and \citet{Hu2017} examined the relative importance of various feedback physics in regulating ISM properties of dwarf galaxies. 
Our work builds upon our current understanding of feedback and galactic chemodynamics while making three notable advances: 1) direct star-by-star modeling, 2) stellar wind feedback from both massive stars and AGB stars, and 3) using an adaptive ray tracing method to follow stellar ionizing radiation. 
%mm2
  We include stellar feedback from stellar winds, core collapse and Type Ia supernovae,    
  photoelectric heating, and ionizing radiation.

Using star-by-star modeling, we capture in more detail the stellar yields from individual stars released over their lifetime. self-consistent set of stellar yields. 
%mm2 
    We track yields from massive star stellar winds, AGB star winds, and supernovae. 
In addition to better capturing how individual metal species enrich the ISM, this allows us to chemically tag individual stars. This ability opens an exciting new channel for testing models of galaxy evolution by leveraging current and ongoing observations probing the detailed 
%mm2
   distributions of
chemical abundances of stars in the Milky Way and Local Group, such as APOGEE and APOGEE2 \citep{APOGEE2010,APOGEE}, the Gaia-ESO survey \citep{Gaia}, and GALAH \citep{GALAH}.
%mm2 
This paper is the first in a series examining in detail the role that the individual components of multi-channel stellar feedback play in galaxy dynamical and chemical evolution. 
%mm2 [this is somehow redundant with the previous paragraph.  I have moved a few things up.]
% In this paper we introduce a new method of feedback physics and galactic chemical enrichment in the cosmological, hydrodynamics code \texttt{Enzo}. For the first time in galaxy-scale simulations, we model star formation down to particles representing individual stars, while simultaneously capturing stellar feedback through stellar winds, core collapse and Type Ia supernovae, photoelectric heating, and ionizing radiation through ray-tracing radiative transfer. In addition, we couple to a self-consistent set of stellar yields
%to track yields from massive star stellar winds, AGB star winds, and supernovae. 
We describe each mode of our multi-channel feedback 
in detail below, and briefly discuss their implementation in an isolated dwarf galaxy simulation.

@@

\section{Methods}
\label{sec:methods}
%mm2 In      summary, 
We produce high-resolution galaxy-scale simulations tracking stars not as SSP's, but as individual stars sampled from an assumed IMF.
This allows us to follow star-by-star variations in feedback physics and stellar yields in detail. To properly model the ISM, we track non-equilibrium, primordial chemistry (including molecular hydrogen) using \texttt{Grackle} \citep{GrackleMethod}, with heating and approximate self-shielding from a metagalactic ultraviolet (UV) background. We 
%mm 
    assume collisional ionization equilibrium for all other elements in order to
use updated \texttt{Cloudy} metal-line cooling tables consistent with our self-shielding approximation (see Appendix ~\ref{appendix:cooling}), and an updated observationally motivated dust model for the low metallicity regimes studied here (Z $\lesssim 0.1$ Z$_{\odot}$). Each star is assigned basic properties (surface gravity, effective temperature, radius, lifetime) from tabulated stellar evolution models, which inform how the stars deposit their feedback. We directly track ionizing radiation from massive stars using an adaptive ray tracing radiative transfer method that includes the effects of radiation pressure. In addition, we follow the optically thin non-ionizing UV radiation from these stars leading to photoelectric heating and Lyman-Werner dissociation of molecular hydrogen. We track the stellar wind feedback and supernovae from these stars, depositing individual metal yields from both. We include AGB wind feedback and yields for lower mass stars, and track these directly as Type Ia supernovae progenitors. We follow yields for 15 individual metal species (C, N, O, Na, Mg, Si, S, Ca, Mn, Fe, Ni, As, Sr, Y, and Ba), chemically tagging each star
%mm
   as it forms
with the associated local gas abundances for each species. In addition, we track a total metal density field which is the sum of all metals, including those not directly tracked. This field is used to inform the heating/cooling physics and determines the metallicity of each star. These methods are discussed in full detail below.
